Parsing Postfix log files

### 2009-01-12 doc/logparser-other-parsers-review.tex
Proofread and corrected the introduction and LMA section of the state of the art review.

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  Ten
parsers have been reviewed as part of the background research for this
project, comparing and contrasting each parser with \parsername{}; the
first parser reviewed is the only previously published research in this
area that the author is aware of.    There are not many previously
developed Postfix log file parsers --- indeed it was quite difficult to
find ten parsers to review for this project --- and the functionality
offered ranges from quite basic to much more mature, depending on the needs
of the author of the parser.  There are some important differences between
\parsername{} and the parsers reviewed here:

@@

@@

    \item None of the parsers reviewed perform the kind of advanced parsing
        and correlation required for this project (described in
        \sectionref{complications})

    \item Only \parsername{} makes it possible to parse new log lines
        without extensive and intrusive modifications to the parser;
        \parsernames{} architecture is described in \sectionref{why
        separate rules, actions, and framework?}.

    \item The parsers reviewed all produce a report of varying complexity
        and detail, whereas \gls{PLP} is designed to be a generic parser,
        extracting data and leaving generation of reports from the data to
        other programs.  Using an \gls{SQL} database simplifies the process
        of generating such reports (discussed in \sectionref{database as
        API}); some simple examples are given in \sectionref{motivation}.

        The parser developed for this project is designed to enable much
        more detailed log file analysis by providing a stable platform for
        subsequent programs to develop upon.

    \item Most of the parsers silently ignore log lines they cannot handle,
        whereas \parsername{} complains loudly about every single log line
        it fails to parse.  The exception is AWStats, which outputs the
        percentage of input lines it was unable to parse, but does not
        output the lines themselves.

    \item A minor difference is that most parsers do not handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic increase
        in parsing time when doing so.  Although this is a minor
        difference, support for reading compressed log files is quite
        helpful, as it dramatically reduces the disk space required to
        store old log files.

    \item Some of the parsers reviewed save the data extracted to a data
        store, but the majority discard all data once they have finished
        generating the report, making historical analysis impossible.

@@

Each of the reviewed parsers was tested with the \numberOFlogFILES{} test
log files described in \sectionref{parser efficiency}.  The data extracted
by \parsername{} is documented in \sectionref{connections table} and
\sectionref{results table}; for convenience that list is repeated here:
server\_ip, server\_hostname, client\_ip, client\_hostname, helo hostname,
queueid, start time, end time, smtp\_code, sender, recipient, size,
message\_id.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \gls{LMA} is quite
different from this parser: it attempts to present correlated data from log
files in a form suitable for a systems administrator to search using the
myriad of standard Unix text processing utilities already available,
producing both a \gls{CSV} file and either a MySQL
(\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \gls{CSV} and Berkeley DB appears to have been
a serious limitation: \gls{CSV} is a very simple format where a record is
stored in a single line, with fields separated by a comma or other
punctuation symbol.  Problems with \gls{CSV} files include the need to
escape separators in the data stored, providing multiple values for a field
(e.g.\ multiple recipients), and adding new fields.  There is no mechanism
to document the fields or the separator, unlike \gls{SQL} databases where
every database includes a schema naming the fields and the type of data
they store (integer, text, timestamp, etc.).

Very little documentation is provided with \gls{LMA}, though some
documentation is available in~\cite{log-mail-analyser}.  Studying the
source code is informative, though this author had difficulty as the
authors of \gls{LMA} wrote in Italian.

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \gls{SQL} databases that store arbitrary tuples.  In \gls{LMA}'s
main table the key is an integer referred to by secondary tables, and the
value is a \gls{CSV} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\gls{IP} tables use the
sender/recipient/date/\gls{IP} as the key, and the value is a \gls{CSV}
list of integers referring to the main table.  This effectively
re-implements \gls{SQL} foreign keys, but without the functionality offered
by even the most basic of \gls{SQL} databases (joins, ordering, searches,
etc.).  It also requires custom code to search on some combination of the
above, though the authors of \gls{LMA} did provide some queries: IP-STORY,
FROM-STORY, DAILY-EMAIL, and DAILY-REJECT\@.  Berkeley DB appears to be the
least useful of the three output formats: it does not provide the
functionality of a basic \gls{SQL} database, nor can it be used with
standard Unix text processing tools.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an \gls{SQL} database, unlike with
Berkeley DB\@; all \gls{SQL} databases embed the schema into the database
and provide commands for displaying it.  Berkeley DB does not embed a
schema, as there is neither requirement nor benefit; it only provides
\textbf{(key, value)} pairs, so any additional structuring of the data is
imposed by the application, thus the application must document this
structure.

The \gls{CSV} record format is not documented, but the output file contains
a comment giving the format:\newline{} \texttt{\# Timestamp|Nome Client|IP
Client|IP Server|From|To|Status|Size} \newline{}\gls{LMA} treats lines
starting with \texttt{\#} as comments, but not all \gls{CSV} parsers will,
as there is no standard for comments in \gls{CSV} files.

\gls{LMA} stores the following date: time and date, client hostname and
\gls{IP} address, server \gls{IP} address, sender and recipient addresses,
\gls{SMTP} code, and size (for accepted mails only); unlike \parsername{}
it does not store the server hostname, helo hostname, queueid, start and
end times, or message id (for accepted mails only).  Handling of multiple
recipients, \gls{SMTP} codes or remote servers\footnote{A single mail may
be sent to multiple remote servers if it was addressed to recipients in
different domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained in the paper; experimental observation shows
that multiple records are added when there are multiple recipients (sadly
the records are not associated or linked in any way), and presumably the
same approach is taken when there are multiple destination servers.

\gls{LMA} requires major changes to the parser code to parse new log lines
or extract additional data.  The code is structured as a long series of
blocks that each handle all lines matching a single regex, so parsing new
log lines requires modifying an existing regex or carefully inserting a new
block in the correct place; extracting extra data will require modifying
multiple blocks, regexes, or both.

\gls{LMA} does not deal with any of the complications discussed in
\sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \gls{LMA} fails
to parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \gls{LMA}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\gls{LMA} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\gls{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  The first three extract \gls{CSV} lines
for the specified client \gls{IP} address, sender address, or date
respectively.  DAILY-REJECT initially failed with an error message from the
Perl interpreter;\footnote{The error message was:
\newline{}\texttt{Undefined subroutine \&main::LIST called at queryDB.pl
line 372.}} after making some corrections to the code it worked, and
extracted the \gls{CSV} lines for the specified day where the \gls{SMTP}
code signifies a rejection.  All of these reports are trivially simple to
produce from the \gls{CSV} file using the standard Unix tool
\texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is merely:

\begin{verbatim}
awk -F\| '$1 ~ /2007-01-26/ && $7 ~ /^[450]|deferred/
    { print $0; print " "; }' lma_output.txt
\end{verbatim}

@@

    \item It outputs an unnecessary line containing only a single space
        after each matching line, to accurately replicate the output of
        DAILY-REJECT\@.

    \item Some records have an \gls{SMTP} code of \textit{0}, though this
        is not a valid \gls{SMTP} code; presumably this is because
        \gls{LMA} did not parse a log line properly.  DAILY-REJECT
        considers this code to be a rejection, so the command above must
        too.

    \item In a small fraction of records the \gls{SMTP} code is
        \textit{deferred\/} (again, not a valid \gls{SMTP} code, and most
        likely because of mis-parsing by \gls{LMA}); it is treated as a
        rejection by DAILY-REJECT, so the equivalent \texttt{awk} command
        must do so, too.

@@

There are two differences between the output from DAILY-REJECT and the
\texttt{awk} command:

@@

    \item \gls{LMA} does not escape or replace the separators used in the
        \gls{CSV} records when they are present in sender or recipient
        addresses, leading to some mangled records; these records are not
        present in the output from \texttt{awk}.  XXX WHY\@?  GIVE AN
        EXAMPLE RECORD\@.

    \item Where there are multiple records in the \gls{CSV} file,
        DAILY-REJECT extracts only one record, whereas the \texttt{awk}
        command above extracts all matching records.

@@

In summary, \gls{LMA} appears to be a proof of concept, written to
demonstrate the point of their paper (that having this information in an
accessible fashion is useful to systems administrators), rather than a
program designed to be useful in a production environment; no consideration
was given to extensibility when writing it.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any data, and it does not
extract any data than is not required for producing its report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not
correlate log lines by queueid or \gls{pid}, and does not need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully parsed the \numberOFlogFILES{} log files
tested with,\footnote{The results it reported were not verified in detail,
but it did not report any errors, and has a very good reputation amongst
Postfix users.} but it is not a suitable base for this project.  It does
not extract the helo hostname, queueid, start and end times, or message id.

@@

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\gls{SMTP} or \gls{ESMTP}).  It parses the
standard syslog\glsadd{syslog} fields at the beginning of the log line, and
extracts any \texttt{key=<value>} pairs occurring after the standard syslog
prologue; these pairs are the fields listed above.  Searches can be based
on the extracted fields or the full text of the log line.  \parsername{}
extracts noticeably more data (client and server \gls{IP} and hostname,
queueid, start and end times, \gls{SMTP} code, and message ID), though it
does not make the full text of the line available (this could be trivially
added if desired, but would greatly increase the size of the resulting
database).  The full power of \gls{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

    \item [Keyfields:] incoming\_host, outgoing\_host, date, hour, sender,
        recipient, defer\_count, delay.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the helo hostname, queueid, \gls{SMTP} code, size of
each accepted mail, and message ID\@.  It extracts some data that
\parsername{} does not: the delay in delivering each mail, and how many
times delivery was deferred for each mail; these could easily be extracted
by \parsername{} if desired.  XXX EXTRACT DELAY AND DEFER ONCE AUTOMATIC
EXTRACTION HAS BEEN FINISHED\@.  Unlike \parsername{}, YAALA does not
maintain separate counters for each restriction; this rules out the
possibility of using the collected data for optimisation, testing or
understanding of restrictions.  YAALA's Postfix parser does not deal with
the complications explained in \sectionref{complications}, though it does
correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: helo hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, and message ID\@.

@@

    \item For mail sent and received: from address, size, and time and
        date.

>>>>

At the start of this project ten Postfix log file parsers were tested, with
the hope of finding a suitable parser to build upon, rather than starting
from scratch.  There are not many Postfix log file parsers available ---
indeed it was quite difficult to find ten parsers to review for this
project --- and the functionality offered ranges from quite basic to much
more mature, depending on the needs of the author of the parser.  None of
those parsers were suitable for this project, so the decision was taken to
write a new parser.  The first parser reviewed is the only previously
published research in this area that the author is aware of; it parses
Postfix log files, but aim of the research is to show that presenting
extracted data in a more easily accessible format is useful to systems
administrators, rather than to improve anti-spam techniques..

The same ten parsers have been reviewed and compared to this project's
finished parser, to show how much effort would have been required to fulfil
the aims and requirements of this project.  It is important to compare and
contrast newly developed algorithms and parsers against those already
available, to accurately judge what improvements, if any, are delivered by
the newcomers.

@@

There are some important differences between \parsername{} and the parsers
reviewed here:

@@

    \item None of the parsers reviewed perform the kind of advanced parsing
        required for this project or deal with the complications described
        in \sectionref{complications}.

    \item Only \parsername{} enables parsing of new log lines without
        extensive and intrusive modifications to the parser; \parsernames{}
        architecture is described in \sectionref{why separate rules,
        actions, and framework?}.

    \item The parsers reviewed all produce a report of varying complexity
        and detail, whereas \gls{PLP} does not; it extracts data and leaves
        generation of reports from the data to other programs.  Using an
        \gls{SQL} database simplifies the process of generating such
        reports (discussed in \sectionref{database as API}); some sample
        queries are given in \sectionref{motivation}.  The parser developed
        for this project is designed to enable much more detailed log file
        analysis by providing a stable platform for subsequent programs to
        develop upon.

    \item Most of the reviewed parsers silently ignore log lines they
        cannot handle, whereas \parsername{} complains loudly about every
        single log line it fails to parse.  The exception is AWStats, which
        outputs the percentage of input lines it was unable to parse, but
        does not output the lines themselves.

    \item A minor difference is that most parsers do not handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic increase
        in parsing time when doing so.  Although this is a minor
        disparity, support for reading compressed log files is quite
        helpful, as it dramatically reduces the disk space required to
        store historical log files.

    \item Some of the parsers reviewed save the extracted data to a data
        store, but the majority discard all data once they have finished
        generating their report, making historical analysis impossible
        without parsing all log files every time.

@@

Each of the reviewed parsers was tested with the \numberOFlogFILES{} test
log files described in \sectionref{parser efficiency}.  The data extracted
by \parsername{} is documented in \sectionref{connections table} and
\sectionref{results table}; for convenience that list is repeated here:
server \gls{IP}, server hostname, client \gls{IP}, client hostname, helo
hostname, queueid, start time, end time, \gls{SMTP} code, sender,
recipient, size, message ID\@.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \gls{LMA} is quite
different from \parsername{}: it attempts to present correlated data from
log files in a form suitable for a systems administrator to search using
the myriad of standard Unix text processing utilities already available.
It produces a \gls{CSV} file and either a MySQL
(\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \gls{CSV} and Berkeley DB appears to have been
a serious limitation: XXX EXTEND\@: LIMITATIONS WILL BE EXPLAINED LATER OR
SOMETHING\@.  Very little documentation is provided with \gls{LMA}, though
some documentation is available in~\cite{log-mail-analyser}.  Studying the
source code is informative, though this author had difficulty as the
authors of \gls{LMA} wrote in Italian.

\gls{CSV} is a very simple format where each record is stored in a single
line, with fields separated by a comma or other punctuation symbol.
Problems with \gls{CSV} files include the need to escape separators in the
data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  There is no standard mechanism to
document the fields or the separator, unlike \gls{SQL} databases where
every database includes a schema naming the fields and the type of data
they store (integer, text, timestamp, etc.).  The \gls{CSV} record format
is not documented, but the output file contains a comment giving the
format:\newline{} \texttt{\# Timestamp|Nome Client|IP Client|IP
Server|From|To|Status|Size} \newline{}\gls{LMA} treats lines starting with
\texttt{\#} as comments, but not all \gls{CSV} parsers will.

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \gls{SQL} databases that store arbitrary tuples.  In \gls{LMA}'s
main table the key is an integer referred to by secondary tables, and the
value is a \gls{CSV} line containing all of the data for that row.  The
secondary by-sender, by-recipient, by-date, and by-\gls{IP} tables use the
sender/recipient/date/\gls{IP} as the key, and the value is a \gls{CSV}
list of integers referring to the main table.  This effectively
re-implements \gls{SQL} foreign keys, but without the functionality offered
by even the most basic of \gls{SQL} databases (joins, ordering, searches,
etc.).  It also requires custom code to search on some combination of the
above, though the authors of \gls{LMA} did provide some queries: IP-STORY,
FROM-STORY, DAILY-EMAIL, and DAILY-REJECT\@.  Berkeley DB appears to be the
least useful of the three output formats: it does not provide the
functionality of a basic \gls{SQL} database, and unlike \gls{CSV} files it
can not be used with standard Unix text processing tools.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an existing \gls{SQL} database, unlike with
Berkeley DB\@; all \gls{SQL} databases embed the schema into the database
and provide commands for displaying it.  Berkeley DB does not embed a
schema, as there is neither requirement nor benefit; it only provides
\textbf{(key, value)} pairs, so any additional structuring of the data is
imposed by the application, thus the application must document this
structure.  MySQL support was not tested because there is no documentation
on the schema required.

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\gls{CSV} output, \gls{LMA} stores the following data: time and date,
client hostname and \gls{IP} address, server \gls{IP} address, sender and
recipient addresses, \gls{SMTP} code, and size (for accepted mails only).
It is unclear which time and date is stored: start time, end time, or
delivery time?  Unlike \parsername{} it does not store the server hostname,
helo hostname, queueid, start and end times, timestamps for each log line,
or message id (for accepted mails only).  Handling of multiple recipients,
\gls{SMTP} codes, or remote servers\footnote{A single mail may be sent to
multiple remote servers if it was addressed to recipients in different
domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained; experimental observation shows that multiple
records are added when there are multiple recipients (sadly the records are
not associated or linked in any way), and presumably the same approach is
taken when there are multiple destination servers.

\gls{LMA} requires major changes to the parser code to parse new log lines
or to extract additional data.  The code is structured as a long series of
blocks that each handle all log lines matching a single regex, so parsing
new log lines requires modifying an existing regex or carefully inserting a
new block in the correct place; extracting extra data will require
modifying multiple blocks, regexes, or both.

\gls{LMA} does not deal with any of the complications discussed in
\sectionref{complications}, except for correlating log lines by queueid;
not correlating log lines by pid means it cannot correlate most rejections.
It does not differentiate between different types of rejections, so it is
not suitable for the purposes of this project; the data about which
restriction caused the rejection is discarded, whereas the main goal of
this project is to retain that data to aid optimisation and evaluation of
anti-spam techniques.  \gls{LMA} fails to parse Postfix log files generated
on Solaris hosts because the fields automatically prepended to each log
line differ from those added on Linux hosts; log files from Solaris hosts
(and possibly other operating systems) thus require preprocessing before
parsing by \gls{LMA}.  Once the preprocessing has been performed on the
\numberOFlogFILES{} test log files \gls{LMA} parses the log files without
complaint, although it produced 32 entries in its output file for every
rejection in the input log file; it also missed some 40\% of delivered
mail.  Once these glaring deficiencies were discovered the author did not
waste any more time checking the results.

\gls{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  These reports search the Berkeley DB files
for matching records: the first three extract \gls{CSV} lines for the
specified client \gls{IP} address, sender address, or date respectively.
DAILY-REJECT initially failed with an error message from the Perl
interpreter;\footnote{The error messages were: \newline{}\texttt{Undefined
subroutine \&main::LIST called at queryDB.pl line
372.}\newline{}\texttt{Undefined subroutine \&main::EXTRACT\_FROM\_DB
called at queryDB.pl line 379.}} after making corrections to the code it
worked, extracting the \gls{CSV} lines for the specified day where the
\gls{SMTP} code signifies a rejection.  All of these reports are trivially
simple to produce from the \gls{CSV} file using the standard Unix tool
\texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is merely:

% perl queryDB.pl -dayreject 2007-01-26 > lma-query

\begin{verbatim}
awk -F\| 'BEGIN { previous = "" };
    $1 ~ /2007-01-26/ && $7 != "250" && $0 != previous 
    { print $0; print " "; previous = $0; }' lma_output.txt
\end{verbatim}

@@

    \item It outputs a line containing only a single space after each
        matching record, to accurately replicate the output of
        DAILY-REJECT\@.

    \item DAILY-REJECT considers all \gls{SMTP} codes except \texttt{250}
        to be rejections; this includes invalid \gls{SMTP} codes such as
        \texttt{0} and \texttt{deferred}, so the awk command does too.
        These invalid \gls{SMTP} codes are most likely present because of
        incorrect parsing by \gls{LMA}.

    \item \gls{LMA} produces 32 output lines in its \gls{CSV} file for
        every single line it should have produced; the command above
        suppresses duplicate sequential lines.

@@

There are some differences between the output from DAILY-REJECT and the
\texttt{awk} command; the author did not spend substantial time attempting
to explain these differences.

@@

    \item The output from DAILY-REJECT is missing some records which are
        present in the \gls{CSV} file; this may be because it uses the
        Berkeley DB files instead, and there may be differences between the
        contents.

    \item Some records output by DAILY-REJECT are truncated: they are
        missing the last | separating fields and the newline following it,
        so the line containing only a single space is concatenated with the
        record.

@@

In summary, \gls{LMA} appears to be a proof of concept, written to
demonstrate the point of their paper (that having this information in an
accessible fashion is useful to systems administrators), rather than a
program designed to be useful in a production environment.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any data, and it does not
extract any data than is not required for producing its report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not
correlate log lines by queueid or \gls{pid}, and does not need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully parsed the \numberOFlogFILES{} log files
tested with,\footnote{The results it reported were not verified in detail,
but it did not report any errors, and has a very good reputation amongst
Postfix users.} but it is not a suitable base for this project.  It does
not extract the helo hostname, queueid, start and end times, timestamps for
each log line, or message id.

@@

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\gls{SMTP} or \gls{ESMTP}).  It parses the
standard syslog\glsadd{syslog} fields at the beginning of the log line, and
extracts any \texttt{key=<value>} pairs occurring after the standard syslog
prologue; these pairs are the fields listed above.  Searches can be based
on the extracted fields or the full text of the log line.  \parsername{}
extracts noticeably more data (client and server \gls{IP} and hostname,
queueid, start and end times, timestamps for each log line, \gls{SMTP}
code, and message ID), though it does not make the full text of the line
available (this could be trivially added if desired, but would greatly
increase the size of the resulting database).  The full power of \gls{SQL}
is available when searching the data extracted by \parsername{}, allowing
the user to search on arbitrarily complicated conditions.

@@

    \item [Keyfields:] incoming\_host, outgoing\_host, date, hour, sender,
        recipient, defer\_count, delay.  Which date and hour are stored is
        not documented: start time, end time, delivery time, or another
        time?

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the helo hostname, queueid, \gls{SMTP} code, size of
each accepted mail, start and end times, timestamps for each log line, and
message ID\@.  It extracts some data that \parsername{} does not: the delay
in delivering each mail, and how many times delivery was deferred for each
mail; these could easily be extracted by \parsername{} if desired.  XXX
EXTRACT DELAY AND DEFER ONCE AUTOMATIC EXTRACTION HAS BEEN FINISHED\@.
Unlike \parsername{}, YAALA does not maintain separate counters for each
restriction; this rules out the possibility of using the collected data for
optimisation, testing or understanding of restrictions.  YAALA's Postfix
parser does not deal with the complications explained in
\sectionref{complications}, though it does correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: helo hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, timestamps of each log line, and message ID\@.

@@

    \item For mail sent and received: from address, size, and time and
        date.  Which date and hour are stored is not documented: start
        time, end time, delivery time, or another time?
